Cycles of the logistic map

### 2013-05-16 doc/pc-ijbc.tex
small change of code
semi-cycle --> half-cycle

\begin{table}[h]\footnotesize
  \tbl{
  Smallest positive $r$ at the onset and bifurcation points
  of the $n$-cycles of the logistic map.
  }
{
\begin{tabular}{llll}
%\begin{tabularx}{\linewidth}{
%  >{\hsize=0.5\hsize\centering\arraybackslash}X
%  >{\hsize=1.6\hsize}X
%  >{\hsize=1.6\hsize}X
%  >{\hsize=0.3\hsize\raggedright\arraybackslash}X
%}
\hline
  $n^\dagger$
& Onset$^\ddagger$
& Bifurcation$^\ddagger$
& \#$^*$ \\
\hline
$1$     & $1.0000000000_1$      &  $3.0000000000_1$       & 1   \\
$2'$    & $3.0000000000_1$      &  $3.4494897428_1$       & 1   \\
$3$     & $3.8284271247_1$      &  $3.8414990075_3$       & 1   \\
$4$     & $3.9601018827_3$      &  $3.9607686524_6$       & 1   \\
$4''$   & $3.4494897428_1$      &  $3.5440903596_6$       & 1   \\
$5$     & $3.7381723753_{11}$   &  $3.7411207566_{15}$    & 3   \\
$6$     & $3.6265531617_{20}$   &  $3.6303887000_{27}$    & 4   \\
$6'$    & $3.8414990075_{3}$    &  $3.8476106612_{27}$    & 1   \\
$7$     & $3.7016407642_{57}$   &  $3.7021549282_{63}$    & 9   \\
$8$     & $3.6621089132_{108}$  &  $3.6624407072_{120}$   & 14  \\
$8'$    & $3.9607686524_{6}$    &  $3.9610986335_{120}$   & 1   \\
$8'''$  & $3.5440903596_{6}$    &  $3.5644072661_{120}$   & 1   \\
$9$     & $3.6871968733_{240}$  &  $3.6872742105_{252}$   & 28  \\
$10$    & $3.6052080669_{472}$  &  $3.6059169323_{495}$   & 48  \\
$10'$   & $3.7411207566_{15}$   &  $3.7425706462_{495}$   & 3   \\
$11$    & $3.6817160194_{1013}$ &  $3.6817266457_{1023}$  & 93  \\
$12$    & $3.5820230011_{1959}$ &  $3.5828117795_{2010}$  & 165 \\
$12'$   & $3.6303887000_{27}$   &  $3.6321857392_{2010}$  & 4   \\
$12''$  & $3.8476106612_{27}$   &  $3.8490363152_{2010}$  & 1   \\
$13$    & $3.6797024578_{4083}$ &  $3.6797038498_{4095}$  & 315 \\
\hline
\multicolumn{4}{p{\linewidth}}{
$^\dagger$
  $\,'$, $\,''$, or $\,'''$ means
    that the cycle is undergoing
    the first, second, or third successive period-doubling, respectively.
} \\
\multicolumn{4}{p{\linewidth}}{
$^\ddagger$
  The subscripts are the degrees of the corresponding minimal polynomial
    of $R = r(r-2)/4$.
} \\
\multicolumn{4}{p{\linewidth}}{
$^*$
  The number of similar cycles.
} \\
\hline
%\end{tabularx}
\end{tabular}
\label{tab:rval}
}
\end{table}

@@

\begin{table}[h]\footnotesize
%\caption{
\tbl{
  Smallest positive $r$ at the onset and bifurcation points
  of the $n$-cycles of the cubic map.
  }
%\begin{center}
{
\begin{tabular}{l l l l}
%\begin{tabularx}{\textwidth}{
%  >{\hsize=0.5\hsize\centering\arraybackslash}X
%  >{\hsize=1.6\hsize}X
%  >{\hsize=1.6\hsize}X
%  >{\hsize=0.3\hsize\raggedright\arraybackslash}X
%}
%\begin{tabular*}{\textwidth}{l l l l c | c l l l l}
\hline
  $n^\dagger$
& Onset$^\ddagger$
& Bifurcation$^\ddagger$
& \#$^*$ \\
\hline
$1$     & $1.0000000000_1$      &  $2.0000000000_1$       & 2   \\
$2'$    & $2.0000000000_1$      &  $2.2360679775_2$       & 2   \\
$3$     & $2.4504409645_{8}$    &  $2.4608286739_{12}$    & 4   \\
$4$     & $2.5478350393_{22}$   &  $2.5488312193_{32}$    & 8   \\
$4''$   & $2.2360679775_{2}$    &  $2.2880317545_{32}$    & 2   \\
$5$     & $2.3939250274_{112}$  &  $2.3957922744_{120}$   & 24  \\
$6$     & $2.3334877526_{304}$  &  $2.3355337580_{336}$   & 56  \\
$6'$    & $2.4608286739_{12}$   &  $2.4657090579_{336}$   & 4   \\
$7$     & $2.3729872678_{1080}$ &  $2.3732727868_{1092}$  & 156 \\
$8$     & $2.3525990555_{3108}$ &  $2.3527637793_{3200}$  & 400 \\
$8'$    & $2.5488312193_{32}$   &  $2.5493247379_{3200}$  & 8   \\
$8'''$  & $2.2880317545_{32}$   &  $2.2992279397_{3200}$  & 2   \\
\hline
\multicolumn{4}{p{\linewidth}}{
$^\dagger$
  $\,'$, $\,''$, or $\,'''$ means that the cycle is undergoing
    the first, second, or third successive period-doubling, respectively.
} \\
\multicolumn{4}{p{\linewidth}}{
$^\ddagger$
  The subscripts are the degrees of the corresponding minimal polynomial.
} \\
\multicolumn{4}{p{\linewidth}}{
$^*$
  The number of similar cycles
    (for $n > 1$, only half of them have positive $r$).
} \\
\hline
\end{tabular}
%\end{tabularx}
%\end{tabular*}
%\end{center}
\label{tab:crval}
}
\end{table}

@@

The characteristic polynomial $A_n\rlam$
  from the determinant equation
  has a degree $N_e(n)$ in $\lambda$.
Again, it encompasses the factors
  from the shorter $d$-cycles ($d|n$),
  and the $n$-cycle polynomial can be obtained
  from \refeq{primfac} by proper substitutions, e.g.,
  $R\rightarrow r$.
%The degree of the polynomial,
The degree in $\lambda$ of $P_n\rlam$
  is given by, after some algebra,
%
%
\begin{equation}
  L_e(n) = \frac{1}{n} \sum_{cd = n} \mu(c)
    \left[
      1 + \odd(c)\frac{3^d-1}{2}
    \right],
\label{eq:cublyndon}
\end{equation}
%
%
For the first few values,
  $L_e(n)$ = 2, 2, 4, 10, 24, 60, 156, 410, \ldots, starting from $n = 1$.
%
This is also the number of the $n$-cycles \cite{hao, hao2}.
Note that, for $n>1$, $x_k$ in half of the cycles
  with negative $r$ are imaginary.
However, in a transformed map
$z_{k+1} = r \, z_k \, (1 - {z_k}^2)$,
which differs from \refeq{cubic} by $x_k = \sqrt{r} z_k$,
$z_k$ in the negative-$r$ cycles are real.
%
%
Following similar steps in \refsec{degR},
we find that
  $\deg_r A_n(r, \lambda) = n N_e(n)$
  and $\deg_r P_n(r, \lambda) = n L_e(n)$.

%\subsection{\label{sec:oddcycle}Odd-cycles}

%
%
\begin{figure}[h]
  \begin{center}
  \begin{minipage}{\linewidth}
        \includegraphics[angle=-90, width=\linewidth]{oddcycle.pdf}
  \end{minipage}%
  \end{center}
  \caption{\label{fig:oddcycle}
  Odd-cycles of the cubic map.}
\end{figure}
%
%

The symmetry $f(-x) = -f(x)$ permits
  a particular \emph{odd-cycle},
  see \reffig{oddcycle} for examples.
If $x_{n+1} = -x_1$ has a solution,
  then $x_1, \ldots, x_{n}, -x_1, \ldots, -x_{n}$
  is a $2n$-cycle, for $x_{2n+1} = - x_{n+1} = x_1$.
%
It follows that a cyclic polynomial
  formed by the $2n$ points
  with an odd degree in $\vxn{2n}$
  is also always zero,
  e.g., $C_1 = x_1 + x_2 + \dots + x_n - x_1 - x_2 - \dots - x_n = 0$.
Accordingly, the characteristic equation \refeq{xr}
  would omit the $n$-odd-cycles from the solution.
The existence of odd-cycles is why we use even, but not odd,
  cyclic polynomials in the solution.

Since the $n$-odd-cycles satisfy a polynomial of lower degree,
  the $n$-cycle polynomial $P_{2n}(\lambda)$
  (omitting the argument $r$)
  can always be factorized.
%
Suppose $\Lambda^o(\vx) \equiv \prod_{k = 1}^{n} f'(x_k)$ in the odd-cycle
 satisfies $P_{n}^o(\lambda^o) = 0$
\big[where $\lambda^o$ is the value of $\Lambda^o(\vx)$,
and $\lambda^o = \pm\sqrt \lambda$\,\big],
then $P^o_{n}(\sqrt \lambda) P^o_{n/2}(-\sqrt \lambda)$
is a factor of $P_{n}(\lambda)$.
%
For example, by solving the $n = 2$ odd-cycle,
  we have $- x_1 = r x_1 - {x_1}^3$, or ${x_1}^2 = r + 1$.
Since $\lambda^o = r - 3 {x_1}^2$,
$P^o_1(\lambda^o) = \lambda^o + 2 r + 3$.
%
The 2-cycle polynomial
  $P_2(\lambda) = -\big[(2r+3)^2 - \lambda\big]\,(\lambda + 2r^2-9)$,
  indeed contains
  $P^o_1(\sqrt \lambda) \, P^o_1(r, -\sqrt \lambda)
   = (\sqrt \lambda + 2 r + 3)(-\sqrt \lambda + 2 r + 3)$
  as a factor.

Finally, we mention that the $n$-cycle polynomial $P_n\rlam$,
  after divided by the contribution from the odd-cycles,
  is an even polynomial of $r$ for an even $n$.
%
This follows from that the transformed variables
  $y_k \equiv (-)^k i x_k$
satisfy
  $y_{k+1} = (-r) y_k - {y_k}^3$.
Thus, there is a one-to-one correspondence
  between the cycles for $r$ and $-r$.
%
Since the mapping applies to an even $m$-odd-cycle,
  hence the corresponding $2m$-cycle,
  $P_n\rlam$ is an even polynomial of $r$
  if $n$ is a multiple of 4.
%
This fact allows one to cut the computating time by half
  in calculating the determinant by Lagrange interpolation
  in these cases.

>>>>

\begin{table}[h]
  %\small
  \tbl{
  Smallest positive $r$ at the onset and bifurcation points
  of the $n$-cycles of the logistic map.
  }
{
\begin{tabular}{llll}
%\begin{tabularx}{\linewidth}{
%  >{\hsize=0.5\hsize\centering\arraybackslash}X
%  >{\hsize=1.6\hsize}X
%  >{\hsize=1.6\hsize}X
%  >{\hsize=0.3\hsize\raggedright\arraybackslash}X
%}
\hline
  $n^\dagger$
& Onset$^\ddagger$
& Bifurcation$^\ddagger$
& \#$^*$ \\
\hline
$1$     & $1.0000000000_1$      &  $3.0000000000_1$       & 1   \\
$2'$    & $3.0000000000_1$      &  $3.4494897428_1$       & 1   \\
$3$     & $3.8284271247_1$      &  $3.8414990075_3$       & 1   \\
$4$     & $3.9601018827_3$      &  $3.9607686524_6$       & 1   \\
$4''$   & $3.4494897428_1$      &  $3.5440903596_6$       & 1   \\
$5$     & $3.7381723753_{11}$   &  $3.7411207566_{15}$    & 3   \\
$6$     & $3.6265531617_{20}$   &  $3.6303887000_{27}$    & 4   \\
$6'$    & $3.8414990075_{3}$    &  $3.8476106612_{27}$    & 1   \\
$7$     & $3.7016407642_{57}$   &  $3.7021549282_{63}$    & 9   \\
$8$     & $3.6621089132_{108}$  &  $3.6624407072_{120}$   & 14  \\
$8'$    & $3.9607686524_{6}$    &  $3.9610986335_{120}$   & 1   \\
$8'''$  & $3.5440903596_{6}$    &  $3.5644072661_{120}$   & 1   \\
$9$     & $3.6871968733_{240}$  &  $3.6872742105_{252}$   & 28  \\
$10$    & $3.6052080669_{472}$  &  $3.6059169323_{495}$   & 48  \\
$10'$   & $3.7411207566_{15}$   &  $3.7425706462_{495}$   & 3   \\
$11$    & $3.6817160194_{1013}$ &  $3.6817266457_{1023}$  & 93  \\
$12$    & $3.5820230011_{1959}$ &  $3.5828117795_{2010}$  & 165 \\
$12'$   & $3.6303887000_{27}$   &  $3.6321857392_{2010}$  & 4   \\
$12''$  & $3.8476106612_{27}$   &  $3.8490363152_{2010}$  & 1   \\
$13$    & $3.6797024578_{4083}$ &  $3.6797038498_{4095}$  & 315 \\
\hline
\multicolumn{4}{p{\linewidth}}{
$^\dagger$
  $\,'$, $\,''$, or $\,'''$ means
    that the cycle is undergoing
    the first, second, or third successive period-doubling, respectively.
} \\
\multicolumn{4}{p{\linewidth}}{
$^\ddagger$
  The subscripts are the degrees of the corresponding minimal polynomial
    of $R = r(r-2)/4$.
} \\
\multicolumn{4}{p{\linewidth}}{
$^*$
  The number of similar cycles.
} \\
\hline
%\end{tabularx}
\end{tabular}
\label{tab:rval}
}
\end{table}

@@

\begin{table}[h]%\footnotesize
%\caption{
\tbl{
  Smallest positive $r$ at the onset and bifurcation points
  of the $n$-cycles of the cubic map.
  }
%\begin{center}
{
\begin{tabular}{l l l l}
%\begin{tabularx}{\textwidth}{
%  >{\hsize=0.5\hsize\centering\arraybackslash}X
%  >{\hsize=1.6\hsize}X
%  >{\hsize=1.6\hsize}X
%  >{\hsize=0.3\hsize\raggedright\arraybackslash}X
%}
%\begin{tabular*}{\textwidth}{l l l l c | c l l l l}
\hline
  $n^\dagger$
& Onset$^\ddagger$
& Bifurcation$^\ddagger$
& \#$^*$ \\
\hline
$1$     & $1.0000000000_1$      &  $2.0000000000_1$       & 2   \\
$2'$    & $2.0000000000_1$      &  $2.2360679775_2$       & 2   \\
$3$     & $2.4504409645_{8}$    &  $2.4608286739_{12}$    & 4   \\
$4$     & $2.5478350393_{22}$   &  $2.5488312193_{32}$    & 8   \\
$4''$   & $2.2360679775_{2}$    &  $2.2880317545_{32}$    & 2   \\
$5$     & $2.3939250274_{112}$  &  $2.3957922744_{120}$   & 24  \\
$6$     & $2.3334877526_{304}$  &  $2.3355337580_{336}$   & 56  \\
$6'$    & $2.4608286739_{12}$   &  $2.4657090579_{336}$   & 4   \\
$7$     & $2.3729872678_{1080}$ &  $2.3732727868_{1092}$  & 156 \\
$8$     & $2.3525990555_{3108}$ &  $2.3527637793_{3200}$  & 400 \\
$8'$    & $2.5488312193_{32}$   &  $2.5493247379_{3200}$  & 8   \\
$8'''$  & $2.2880317545_{32}$   &  $2.2992279397_{3200}$  & 2   \\
\hline
\multicolumn{4}{p{\linewidth}}{
$^\dagger$
  $\,'$, $\,''$, or $\,'''$ means that the cycle is undergoing
    the first, second, or third successive period-doubling, respectively.
} \\
\multicolumn{4}{p{\linewidth}}{
$^\ddagger$
  The subscripts are the degrees of the corresponding minimal polynomial.
} \\
\multicolumn{4}{p{\linewidth}}{
$^*$
  The number of similar cycles
    (for $n > 1$, only half of them have positive $r$).
} \\
\hline
\end{tabular}
%\end{tabularx}
%\end{tabular*}
%\end{center}
\label{tab:crval}
}
\end{table}

@@

The characteristic polynomial $A_n\rlam$
  from the determinant equation
  has a degree $N_e(n)$ in $\lambda$.
Again, it encompasses the factors
  from the shorter $d$-cycles ($d|n$),
  and the $n$-cycle polynomial can be obtained
  from \refeq{primfac} by proper substitutions, e.g.,
  $R\rightarrow r$.
%The degree of the polynomial,
The degree in $\lambda$ of $P_n\rlam$
  is given by, after some algebra,
%
%
\begin{equation}
  L_e(n) = \delta_{n, 1} +
    \frac{1}{n} \sum_{cd = n} \mu(c) \odd(c)\frac{3^d-1}{2},
\label{eq:cublyndon}
\end{equation}
%
%
For the first few values,
  $L_e(n)$ = 2, 2, 4, 10, 24, 60, 156, 410, \ldots, starting from $n = 1$.
%
This is also the number of the $n$-cycles \cite{hao, hao2}.
Note that, for $n>1$, $x_k$ in half of the cycles
  with negative $r$ are imaginary.
However, in a transformed map
$z_{k+1} = r \, z_k \, (1 - {z_k}^2)$,
which differs from \refeq{cubic} by $x_k = \sqrt{r} z_k$,
$z_k$ in the negative-$r$ cycles are real.
%
%
Following similar steps in \refsec{degR},
we find that
  $\deg_r A_n(r, \lambda) = n N_e(n)$
  and $\deg_r P_n(r, \lambda) = n L_e(n)$.

%\subsection{\label{sec:halfcycle}Odd-cycles}

%
%
\begin{figure}[h]
  \begin{center}
  \begin{minipage}{\linewidth}
        \includegraphics[angle=-90, width=\linewidth]{halfcycle.pdf}
  \end{minipage}%
  \end{center}
  \caption{\label{fig:halfcycle}
  Odd-cycles of the cubic map.}
\end{figure}
%
%

The symmetry $f(-x) = -f(x)$ permits
  a particular \emph{half-cycle},
  see \reffig{halfcycle} for examples.
If $x_{n+1} = -x_1$ has a solution,
  then $x_1, \ldots, x_{n}, -x_1, \ldots, -x_{n}$
  is a $2n$-cycle, for $x_{2n+1} = - x_{n+1} = x_1$.
%
It follows that a cyclic polynomial
  formed by the $2n$ points
  with an odd degree in $\vxn{2n}$
  is also always zero,
  e.g., $C_1 = x_1 + x_2 + \dots + x_n - x_1 - x_2 - \dots - x_n = 0$.
Accordingly, the characteristic equation \refeq{xr}
  would omit the $n$-half-cycles from the solution.
The existence of half-cycles is why we use even, but not odd,
  cyclic polynomials in the solution.

Since the $n$-half-cycles satisfy a polynomial of lower degree,
  the $n$-cycle polynomial $P_{2n}(\lambda)$
  (omitting the argument $r$)
  can always be factorized.
%
Suppose $\Lambda^o(\vx) \equiv \prod_{k = 1}^{n} f'(x_k)$ in the half-cycle
 satisfies $P_{n}^o(\lambda^o) = 0$
\big[where $\lambda^o$ is the value of $\Lambda^o(\vx)$,
and $\lambda^o = \pm\sqrt \lambda$\,\big],
then $P^o_{n}(\sqrt \lambda) P^o_{n/2}(-\sqrt \lambda)$
is a factor of $P_{n}(\lambda)$.
%
For example, by solving the $n = 2$ half-cycle,
  we have $- x_1 = r x_1 - {x_1}^3$, or ${x_1}^2 = r + 1$.
Since $\lambda^o = r - 3 {x_1}^2$,
$P^o_1(\lambda^o) = \lambda^o + 2 r + 3$.
%
The 2-cycle polynomial
  $P_2(\lambda) = -\big[(2r+3)^2 - \lambda\big]\,(\lambda + 2r^2-9)$,
  indeed contains
  $P^o_1(\sqrt \lambda) \, P^o_1(r, -\sqrt \lambda)
   = (\sqrt \lambda + 2 r + 3)(-\sqrt \lambda + 2 r + 3)$
  as a factor.

Finally, we mention that the $n$-cycle polynomial $P_n\rlam$,
  after divided by the contribution from the half-cycles,
  is an even polynomial of $r$ for an even $n$.
%
This follows from that the transformed variables
  $y_k \equiv (-)^k i x_k$
satisfy
  $y_{k+1} = (-r) y_k - {y_k}^3$.
Thus, there is a one-to-one correspondence
  between the cycles for $r$ and $-r$.
%
Since the mapping applies to an even $m$-half-cycle,
  hence the corresponding $2m$-cycle,
  $P_n\rlam$ is an even polynomial of $r$
  if $n$ is a multiple of 4.
%
This fact allows one to cut the computating time by half
  in calculating the determinant by Lagrange interpolation
  in these cases.
